A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-06-24 sections/algorithm.tex
typos
shrink to fit in 27 pages

\paragraph{Example} The derivation of the subtyping judgment
$(1\to 1)\to 1 \le (\all 1\to 1) \to 1$:
\begin{gather*}
\begin{aligned}
  & \nil\vdash (1\to 1)\to 1 \le (\all 1 \to 1) \to 1\\
\rrule{7}  & \nil \Vdash 1 \le 1 \Vdash \all 1 \to 1 \le 1 \to 1\\
\rrule{8}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1 \le 1 \to 1\\
\rrule{7}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \le 1 \Vdash 1 \le 1\\
\rrule{4} & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \le 1\\
\rrule{4} & \nil \Vdash 1 \le 1 ,\al\\
\rrule{2} & \nil \Vdash 1 \le 1\\
\rrule{4} & \nil
\end{aligned}
\end{gather*}
First, the subtyping of two function types is split into two judgments by Rule 7:
a covariant subtyping on the return type and a contravariant subtyping on the argument type.
Then we apply Rule 8 to reduce the $\forall$ quantifier on the left side.
The rule introduces an existential variable $\al$ to the context (even though 
the type $\all 1 \to 1$ does not actually refer to the quantified type
variable $a$).
In the following 3 steps we satisfy the judgment $1 \to 1 \le 1 \to 1$ by Rules 7, 4 and 4.

@@

After that, Rule 28 changes the application inference judgment to
a check of the argument against the input type $\al$ and returns the output type $\al$.
Checking $()$ against the existential variable $\al$ solves $\al$ to the unit type $1$
through Rules 18, 24 and 16.
Immediately after $\al$ is solved, the algorithm replaces every occurrence of $\al$ with $1$.
Therefore the worklist remains a simple subtyping judgment $1 \le 1$, which is finished off by Rule 4.
Finally we get an empty worklist, indicating the success of the whole derivation.

>>>>

\paragraph{Example} Consider the derivation of the subtyping judgment
$(1\to 1)\to 1 \le (\all 1\to 1) \to 1$:
\begin{gather*}
\begin{aligned}
  & \nil\vdash (1\to 1)\to 1 \le (\all 1 \to 1) \to 1\\
\rrule{7}  & \nil \Vdash 1 \le 1 \Vdash \all 1 \to 1 \le 1 \to 1\\
\rrule{8}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \to 1 \le 1 \to 1\\
\rrule{7}  & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \le 1 \Vdash 1 \le 1\\
\rrule{4} & \nil \Vdash 1 \le 1 ,\al \Vdash 1 \le 1\\
\rrule{4} & \nil \Vdash 1 \le 1 ,\al\\
\rrule{2} & \nil \Vdash 1 \le 1\\
\rrule{4} & \nil
\end{aligned}
\end{gather*}
First, the subtyping of two function types is split into two judgments by Rule 7:
a covariant subtyping on the return type and a contravariant subtyping on the argument type.
Then we apply Rule 8 to reduce the $\forall$ quantifier on the left side.
The rule introduces an existential variable $\al$ to the context (even though 
the type $\all 1 \to 1$ does not actually refer to the quantified type
variable $a$).
In the following 3 steps we satisfy the judgment $1 \to 1 \le 1 \to 1$ by Rules 7, 4 and 4.

@@

After that, Rule 28 changes the application inference judgment to
a check of the argument against the input type $\al$ and returns the output type $\al$.
Checking $()$ against the existential variable $\al$ solves $\al$ to the unit type $1$
through Rules 18, 24 and 16.
Immediately after $\al$ is solved, the algorithm replaces every occurrence of $\al$ with $1$.
Therefore the worklist remains $1 \le 1$, which is finished off by Rule 4.
Finally, the empty worklist indicates the success of the whole derivation.
